\phantomsection
\subsection*{III.5—Comparative Synthesis: Binding, Punishment, Healing}
\addcontentsline{toc}{subsection}{III.5—Comparative Synthesis}
\label{ssec:iii-synthesis}

Hebrew Scripture, Christian Testament, patristic-medieval theology, and early Buddhism
converge on the paradox that fulfillment can be judgment---yet they diverge profoundly in
ontology, diagnosis, and remedy.

\textbf{What binds?} Biblical tradition names the bondage as \emph{sin}: love disordered
toward finite goods as if ultimate. Psalm 106 and 1 Samuel 8 show Israel getting what it wants
and discovering the want itself was misdirected. Augustine systematizes this as
\emph{cupiditas} (possessive desire) versus \emph{caritas} (rightly ordered love)
\gbparencite{AugustineConfessions1998}. Aquinas locates the disorder in \emph{aversio a Deo}:
turning from the infinite Good toward creatures absolutized \gbparencite{AquinasST1947}. The
problem is not desire \emph{per se} but its target and manner.

Buddhist tradition names the bondage as \emph{taṇhā-upādāna}: craving and clinging that arise
from ignorance of impermanence and non-self. The \emph{Dhammapada} and \emph{Saṃyutta Nikāya}
diagnose suffering as endogenous to grasping \parencite{BodhiSN2000}. Nāgārjuna adds that
craving reifies phenomena as inherently existent, setting up the wound when their emptiness is
disclosed \parencite{NagarjunaMMK2013}. The problem is not misdirection but
\emph{direction-as-such}---any attachment to conditioned phenomena perpetuates becoming.

\textbf{What punishes?} In Biblical-patristic idiom, the will is ``handed over'' to its object
(Romans 1:24), or fulfillment brings ``leanness of soul'' (Psalm 106:15). Augustine writes
that disordered affection is its own punishment; Aquinas that turning from God entails
privation. The punishment is not external retribution but the intrinsic failure of finite goods
to bear infinite demand. The mechanism is teleological: a will aimed wrongly cannot reach its
true end, and the mismatch is experienced as emptiness, fragmentation, or despair.

In Buddhist idiom, gratification feeds the next link in dependent origination. Craving
conditions clinging, clinging conditions becoming, becoming conditions birth, birth conditions
aging-and-death (SN 12.2). The punishment is causal and impersonal: no judge assigns penalty,
but the structure of conditioned arising guarantees that satisfied craving generates new
suffering. The Fire Sermon's imagery---sense faculties ``burning'' with lust, hate,
delusion---shows that the fire is not fuel-dependent; it \emph{is} the craving itself, so
``feeding'' it intensifies rather than extinguishes it.

Both traditions understand Coppola's sampan episode identically in form but differently in
ground: the mission ``succeeds,'' the unarmed civilians are killed, the crew's affect goes
hollow. Biblical interpretation reads this as disordered desire (securing the mission as
ultimate) exposed through its own success. Buddhist interpretation reads it as craving
(security, mission-completion) perpetuating suffering through another link in the chain. The
phenomenology is the same; the metaphysics differ.

\textbf{What heals?} Biblical-patristic tradition proposes grace re-ordering love. Augustine
insists that the will, enslaved to \emph{cupiditas}, cannot free itself but must be healed by
participation in divine love \gbparencite{AugustineCity2003}. Aquinas agrees: only when desire
is rightly ordered toward God as ultimate end can finite goods be enjoyed without privation
\gbparencite{AquinasST1947}. Gregory of Nyssa's \emph{epektasis} suggests that infinite desire
directed toward the infinite God becomes non-possessive ascent rather than grasping
\gbparencite{GregoryMoses1978}. Maimonides offers a rationalist variant: intellectual
discipline reorders the soul so that appetite serves rather than obscures the vision of God
\gbparencite{MaimonidesGuide1963}. The remedy is \emph{reorientation}: not ceasing to desire
but desiring rightly.

Buddhist tradition proposes \emph{cessation}. The Noble Eightfold Path cultivates
disenchantment with conditioned phenomena, leading to the cooling of craving and liberation
from the cycle \parencite{Rahula1959}; \parencite{Gethin1998}. Walpola Rahula writes: ``The
cessation of dukkha is the cessation of taṇhā.'' There is no reorientation toward a higher
Good but a progressive unbinding from all objects of attachment. Śāntideva's
\emph{Bodhicaryāvatāra} complicates this by redirecting compassion toward all sentient
beings---a kind of ethical desire---but even here the goal is eventual liberation rather than
eternal participation \parencite{SantidevaBCA1995}.

The two paths are not easily reconciled. Biblical-patristic thought insists that the soul is
made for God and only infinite Good can satisfy; desire rightly ordered ascends eternally
without exhaustion. Early Buddhism insists that any conditioned object, even ``God,'' remains
within the cycle and thus subject to dukkha; only cessation of craving brings peace. One
infinitizes desire toward the Absolute; the other extinguishes grasping altogether.

Yet both agree on the diagnostic: finite fulfillment, sought possessively, punishes. The
administrative mission that Willard receives promises clarity, purpose, even redemption. It
delivers exposure: of complicity (Biblical reading) or of samsaric repetition (Buddhist
reading). ``Everyone gets everything he wants''---the structure is universal. ``For my sins I
got one''---the judgment is intrinsic. The film offers no remedy, only the vision of
fulfillment as wound. Whether healing requires reordered love or cessation of craving remains
theologically contested. What the river journey confirms is that getting what one wants, when
the wanting is disordered or grasping, does not liberate. It binds---and the binding is the
punishment.
