\phantomsection
\section*{Appendix A: On AI Authorship and Process}
\addcontentsline{toc}{section}{Appendix A: On AI Authorship and Process}
\label{app:ai-authorship}

Other than this appendix, the essay is entirely written by AI using OpenAI's ChatGPT
\parencite{OpenAIChatGPT2024} and Cursor \parencite{CursorAI2024}. It started as a question
to figure out the meaning of Willard's line, however it seemed the answer could be complicated
and was approaching an essay, so the human author went forward with asking the AI to expand
it. This essay does not exist for me to convince others of my thesis, but for AI to tell me
exactly what the film meant by viewing it through a lense of theological and philosophical
interpretations.

The human author acted mostly like a teacher---I gave the essay prompt and topic, worked
with the AI student on developing an outline, and kept sending it back with criticism and
requests for revisions. The human author has not ever read a slight majority of the works
cited, but is hopeful that asking for specific quotes and citations to actual page numbers in
actual works is keeping AI a bit more honest, and I am comforted that the explanations
regarding works I am familiar with seem to be reasonable. The full OpenAI interaction log and
Cursor interaction logs are in the accompanying
\href{https://github.com/jmalicki/apocalypse-now-essay/tree/main/interaction-logs}{git
	repository}, where you can see the work that the human author did to produce the essay.

This is not an attempt at scholarly look, but just me having fun.
I just wanted to understand the film's deeper
meaning, and I asked AI because Google did not give me the answers.  Everyone gets everything
he wants.  I wanted to read an essay going deep on what philosophers through time would feel
about the thesis of the film, and for my sins, AI gave me one.
