
\phantomsection
\section*{V. Colonial Modernity and the Critique of Domination}
\addcontentsline{toc}{section}{V. Colonial Modernity and Critique of Domination}
\label{sec:v-colonial-modernity}

Having traced theological, philosophical, and psychological interpretations of Willard's line, 
we now turn to critical theory's historicization of the will's structure. Coppola's film 
locates Conrad's private empire within imperial modernity—a world where fulfillment is 
bureaucratized, representation circulates power, and the will internalizes discipline. The 
twentieth century's thinkers show how ``getting what one wants'' exposes not only metaphysical 
structures but historical complicity.

\subsection*{1. From Conrad to Coppola: Bureaucracy as a Technology of Will}
\label{ssec:1-from-conrad-to-coppola-bureaucracy-as-a-technology-of-will}
Coppola's adaptation transposes Conrad's private empire into a military bureaucracy that 
routinizes transgression. Willard's orders are typed, briefed, and accompanied by dossiers; 
Kurtz's poetry is replaced by radio logs and classified memoranda. The mission is thus an 
artifact of files, not of metaphysical vocation. In Willard's terms, ``for my sins they gave me 
one'': the administrative system internalizes the will's desire for a task and returns it as 
obligation \parencite{CoppolaApocalypse1979}. The film's mise-en-scène---the air cavalry's 
Wagnerian assault, the Playboy USO show, the bridge to nowhere---presents modern fulfillment as 
spectacle: an economy of images where desire circulates as command and entertainment at once.

Adorno and Horkheimer called this dialectic in advance: the Enlightenment's will to demystify 
nature turns into domination of humans by technical reason \parencite{AdornoHorkheimer2002}. 
The same rationality that frees us from myth installs the world as object of control. In 
\textit{Apocalypse Now}, the mission is in this sense an \emph{Enlightenment object}: planned, 
staged, justified. Its fulfillment exposes the subject as functionary. Arendt's analysis of 
modern totalitarianism clarifies the disjunction between action and responsibility here: systems 
generate outcomes that no single agent intends, but which implicate every actor caught within 
them \parencite{ArendtOrigins1973}. Willard's errand-boy status embodies this structure of 
dispersed agency.

\subsection*{2. The Spirit of Domination: Work, Discipline, and Representation}
\label{ssec:2-the-spirit-of-domination-work-discipline-and-representation}
Weber’s thesis on the Protestant ethic connects transcendent assurance to immanent compulsion: the anxiety of salvation is displaced into the worldly signs of vocation, productivity, and discipline \parencite{WeberProtestant2002}. In Willard’s world, vocation has lost its soteriological frame; the residue is compulsion alone. The river journey is a pilgrimage without grace: a labor of proof that can never culminate. Foucault’s analysis of modern power makes the continuity plain: discipline produces subjects by normalizing their bodies and perceptions \parencite{FoucaultDiscipline1995}. Willard’s training, files, and surveillance are not contingent backdrops; they \emph{are} the conditions under which a ``mission'' can be willed, received, and fulfilled. The will internalizes the gaze.

If domination requires a world to dominate, representation supplies it. Said shows how the Orient is constructed as an object of knowledge that authenticates Western authority \parencite{SaidOrientalism1978}. \textit{Apocalypse Now} multiplies such representations: the radio’s ``psyops'' patter, military briefings, newspaper clippings, and narration. The hilltop massacre under flares is not only an event but an image of an event; it exists to be seen. Benjamin’s ``Angel of History'' looks back not upon progress but upon ``a single catastrophe which keeps piling wreckage upon wreckage'' \parencite{BenjaminTheses1969}. The film literalizes this gaze: moving forward upriver is moving back into debris. Fulfillment of the mission produces a tableau of ruins through which the angel is blown.

\subsection*{3. Recognition, Violence, and the Will to Purity}
\label{ssec:3-recognition-violence-and-the-will-to-purity}
Fanon reinterprets Hegelian recognition within colonial relation: the colonized subject meets the colonizer’s will to purity as violence; the only available agency appears as counter-violence \parencite{FanonWretched2004}. Kurtz absolutizes this logic. His ``methods'' are ``pure'' because they purge ambivalence. He wants an end to contradiction; he wants an act that would finally coincide with intent. To get this is to erase the human. Willard confronts not only a man but the fantasy of unmediated will. Here the line between Section III’s existentialism and Section IV’s historicity thins: the metaphysics of will finds its historical instrument in colonial modernity. What the will wants (sovereignty) appears in the world as the right to decide life and death.

\subsection*{4. Fulfillment as Exposure: Modernity's Mirror}
\label{ssec:4-fulfillment-as-exposure-modernity-s-mirror}
Read in this frame, Willard’s opening line is not merely mordant wit but a summary of the century’s critique. ``Everyone gets everything he wants''---because modern institutions exist to circulate wants as functions, and because representation manufactures the worlds those wants require. ``For my sins I got one''---because the system returns desire as assignment, and the assignment reveals desire’s complicity with domination. Conrad supplies the form (a journey into the center where fulfillment collapses into horror). Critical theory supplies the terms (reason as domination, vocation as compulsion, representation as power). Coppola supplies the image: the fulfilled mission as an illuminated ruin.

What Willard learns upriver is what Conrad, Benjamin, and Fanon teach in theory: fulfillment is not closure but exposure. The prize of the modern will is to see itself in the world it has made.
