\phantomsection
\subsection*{VI.1—Freud: The Death Drive, Repetition Compulsion, and Civilization's Discontents}
\addcontentsline{toc}{subsection}{VI.1—Freud: Death Drive and Repetition}
\label{ssec:vi-freud}

Sigmund Freud's later work confronts phenomena that unsettle his earlier pleasure-principle 
model: why do neurotics compulsively repeat traumatic experiences? Why do patients resist cure? 
In \textit{Beyond the Pleasure Principle} (1920), Freud hypothesizes a ``death drive'' 
(\emph{Todestrieb})---a tendency in all organic life to ``restore an earlier state of things,'' 
ultimately the inanimate \parencite[p.~38]{FreudBeyond1955}. This drive does not seek pleasure 
but discharge, not satisfaction but return to zero. The ``daemonic'' character of repetition 
``overrides the pleasure principle'' \parencite[p.~22]{FreudBeyond1955}: the subject 
re-enacts precisely what wounded him, as if compelled to master through repetition what he 
could not master in the original experience.

Willard's opening confession---``I wanted a mission''---reads, in this frame, as a repetition 
compulsion. He has already completed missions; they did not satisfy. Yet he wants another, as 
if the next iteration might finally discharge the tension the previous ones installed. The 
upriver journey literalizes this: each checkpoint repeats the pattern (encounter, violence, 
continuation), and none resolves. The Do Lung Bridge, rebuilt nightly only to be destroyed, is 
Freud's fort-da game writ large---the child throws the spool away and reels it back, mastering 
absence through symbolic repetition \parencite[pp.~14--16]{FreudBeyond1955}. But mastery never 
arrives; the compulsion only intensifies.

In \textit{Civilization and Its Discontents} (1930), Freud reframes the contradiction 
socially: civilization requires renunciation of instinct, and the price is guilt. The more 
successfully culture sublimates aggression into work and law, the more the leftover aggression 
turns inward as superego \parencite[pp.~70--97]{FreudCivilization1961}. This is why ``getting 
what one wants'' (order, justice, security) feels punitive: the very achievements of 
civilization intensify the psychic burden. Willard's mission is both sublimation (redirecting 
violent impulse into ``lawful'' assassination) and the return of what was sublimated (he must 
kill to preserve the order that forbids killing). The guilt is structural, not contingent.

Freud also distinguishes Eros (binding, connection) from Thanatos (unbinding, aggression). 
Civilization is the battleground where these forces clash \parencite[pp.~81--92]{FreudCivilization1961}. 
The mission-form channels Thanatos under the sign of Eros: violence is presented as care 
(``terminate to restore order''), destruction as construction. When the mission is fulfilled, 
the Eros-claim collapses, leaving only the Thanatos that was operative all along. ``For my 
sins I got one'' is the moment the sublimation fails and the drive's true orientation is 
revealed. Fulfillment punishes because it strips the Eros-mask from Thanatos.

Finally, Freud insists that the unconscious knows no negation and no time: repressed content 
persists, undischarged, awaiting return \parencite[pp.~166--171]{FreudBeyond1955}. Willard's 
voiceover has this timeless, circular quality---he narrates the past as if it were still 
present, the mission as if it were still ahead. The journey cannot be completed 
\emph{psychically} even when it is completed procedurally. The death drive ensures that 
satisfaction is always deferred, always returned as renewed lack. ``Everyone gets everything 
he wants'' becomes, in Freud's grammar, Everyone receives the object, but the drive persists. 
The wanting was never about the object.

